\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-09 paper.tex
hyphenate subtle "domain specific"
edits in response to Josh comments

Automatically inducing programs from examples is a long-standing goal
of artificial intelligence. Recent work has successfully used
symbolic search techniques (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}), neural networks trained from
a corpus of examples (e.g., RobustFill:~\cite{devlin2017robustfill},
DeepCoder:~\cite{balog2016deepcoder}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search~\cite{ngds}) to
synthesize programs for task domains such as string transformations,
list processing, and robot navigation and planning. However, all these
approaches -- symbolic, neural and neural-symbolic -- rely upon a
hand-engineered \emph{Domain-Specific Language} (DSL). The DSL is an
inventory of restricted programming primitives, encoding domain
specific knowledge about the space of programs. In
practice
we often have only a few input/output examples for each
program to be induced, and thus success often hinges on having a good
DSL which
provides a crucial inductive bias for what would otherwise be an
unconstrained search through the space of all computable functions.
Here we ask, to what extent can we dispense with such highly
hand-engineered domain-specific languages?

@@

\begin{algorithm}[tb]
   \caption{The \system Algorithm}
   \label{mainAlgorithm}
   \begin{algorithmic}
     \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
     \STATE \textbf{Hyperparameters:} Maximum frontier size $F$
     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
     \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
     \FOR{$i=1$ {\bfseries to} $I$}
     %     \FOR{$x:\tau\in X$}
     \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enumerate}(\mathcal{D},\theta,F)\text{ if }\probability[x|p] > 0\}$
 %    \ENDFOR
     \STATE $q(\cdot )\gets \text{train recognition model to maximize }\mathcal{L}_{\text{RM}}$
     \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enumerate}(\mathcal{D},q(x),F)\text{ if }\probability[x|p] > 0\}$
     \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$
     %% \STATE Define $Q_x(z) \propto \begin{cases}
     %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
     %%   0&x\not \in \mathcal{F}_x
     %% \end{cases}
      \ENDFOR
 \STATE \textbf{return} $\mathcal{D},\theta,q$
\end{algorithmic}
\end{algorithm}

@@

We compare  with four baselines on  held-out tasks:
\\\noindent \textbf{Ours (no $q$)}, which lesions the recognition model, $q$.
\\\noindent \textbf{RF/DC}, which holds $(\mathcal{D},\theta)$ fixed and learns
a recognition model only from samples from the fixed generative model.
This is equivalent to our algorithm with $\lambda = \infty$ (see Sec.~\ref{grammarInductionSection})
and $\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{HM}}$ (see Sec.~\ref{recognitionSection}).
We call this baseline RF/DC because
this setup is closest to how RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
We can not compare directly with these systems,
because they are engineered for one specific domain, and 
do not have publicly available code and datasets.
\\\noindent \textbf{PCFG}, which lesions the recognition model, learns $\theta$, and fixes $\mathcal{D}$.
This is equivalent to our algorithm with $q(x) = \theta$ and $\lambda = \infty$,
and is like learning the parameters of a PCFG while not learning any of the structure.
\\\noindent \textbf{Enum}, which does no learning and just enumerates a frontier.
This is equivalent to our first wake cycle.

@@

\begin{table}
%% This radically shrinks the table
\tabcolsep=4pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{llllll}
  \toprule& Ours& \begin{tabular}{c}
    Ours\\(no $q$)
    \end{tabular}& RF/DC & PCFG & Enum

>>>>

Automatically inducing programs from examples is a long-standing goal
of artificial intelligence. Recent work has successfully used
symbolic search techniques (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}), neural networks trained from
a corpus of examples (e.g., RobustFill:~\cite{devlin2017robustfill},
DeepCoder:~\cite{balog2016deepcoder}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search~\cite{ngds}) to
synthesize programs for task domains such as string transformations,
list processing, and robot navigation and planning. However, all these
approaches -- symbolic, neural and neural-symbolic -- rely upon a
hand-engineered \emph{Domain-Specific Language} (DSL). The DSL is an
inventory of restricted programming primitives, encoding domain-specific
knowledge about the space of programs. In
practice
we often have only a few input/output examples for each
program to be induced, and thus success often hinges on having a good
DSL which
provides a crucial inductive bias for what would otherwise be an
unconstrained search through the space of all computable functions.
Here we ask, to what extent can we dispense with such highly
hand-engineered domain-specific languages?

@@

\begin{algorithm}[tb]
   \caption{The \system Algorithm}
   \label{mainAlgorithm}
   \begin{algorithmic}
     \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
     \STATE \textbf{Hyperparameters:} Maximum frontier size $F$
     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
     \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
     \FOR{$i=1$ {\bfseries to} $I$}
     %     \FOR{$x:\tau\in X$}
     \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in \text{enum}(\mathcal{D},\theta,F)\text{ if }\probability[x|p] > 0\}$ \footnotesize{(\textbf{Wake})}
     \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ (\footnotesize{\textbf{Sleep-R}})
     \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in \text{enum}(\mathcal{D},q(x),F)\text{ if }\probability[x|p] > 0\}$ \footnotesize{\textbf{(Wake)}}

     \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$   \footnotesize{\textbf{(Sleep-G)}}
     %% \STATE Define $Q_x(z) \propto \begin{cases}
     %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
     %%   0&x\not \in \mathcal{F}_x
     %% \end{cases}
      \ENDFOR
 \STATE \textbf{return} $\mathcal{D},\theta,q$
\end{algorithmic}
\end{algorithm}

@@

We compare  with four baselines on  held-out tasks:
\\\noindent \textbf{Ours (no NN)}, which lesions the recognition model, $q$.
\\\noindent \textbf{RF/DC}, which holds $(\mathcal{D},\theta)$ fixed and learns
a recognition model only from samples from the fixed generative model.
This is equivalent to our algorithm with $\lambda = \infty$ (see Sec.~\ref{grammarInductionSection})
and $\mathcal{L}_{\text{RM}} = \mathcal{L}_{\text{HM}}$ (see Sec.~\ref{recognitionSection}).
We call this baseline RF/DC because
this setup is closest to how RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
We can not compare directly with these systems,
because they are engineered for one specific domain, and 
do not have publicly available code and datasets.
\\\noindent \textbf{PCFG}, which lesions the recognition model, learns $\theta$, and fixes $\mathcal{D}$.
This is equivalent to our algorithm with $q(x) = \theta$ and $\lambda = \infty$,
and is like learning the parameters of a PCFG while not learning any of the structure.
\\\noindent \textbf{Enum}, which does no learning and just enumerates a frontier.
This is equivalent to our first wake cycle.

@@

\begin{table}
%% This radically shrinks the table
\tabcolsep=4pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{llllll}
  \toprule& Ours& \begin{tabular}{c}
    Ours\\(no NN)
    \end{tabular}& RF/DC & PCFG & Enum
